% The abstract should briefly summarize the contents of the paper in
% 150--250 words.
% \keywords{First keyword  \and Second keyword \and Another keyword.}
The acquisition of multi-view datasets, that is containing multiple brain imaging modalities with co-occurring biological and clinical variables, is a very common practice in neuroimaging dementia studies.
Thanks to big data technologies, researchers can easily obtain multi-view data from different studies simultaneously,
although the massive sample size, high dimensionality, and missing data introduce unique computational and statistical challenges when using machine learning and artificial intelligence techniques for the analysis.
Indeed, the joint modeling of different neuroimaging datasets requires the ability of consistently analyzing high-dimensional heterogeneous information in presence of often non-overlapping sets of views, or modalities (e.g. clinical scores, imaging data, or biological measurements) across data samples.
This problem requires to simultaneously deal with random and non-random missing data specific to each dataset, under the form of missing measurements corresponding to a given view.
%
We propose here a generative latent-variable model of the common variability across datasets, based on the estimation of a shared latent representation across heterogeneous views.
Our formulation allows to retrieve a consistent latent representation common to all data-types and datasets, even if datasets share only partially overlapping subsets of views, in a computationally efficient and scalable manner.
%
Simulations on synthetic data show that our method is able to identify a common latent representation of multi-feature datasets, even when the overlap of views in the feature-sets across datasets is minimal.
% The resulting latent representation is generally equivalent to one obtained with a model trained with no missing data-types.
%
When analyzing multi-modal neuroimaging and clinical data from independent dementia studies jointly, our model is able to mitigate the absence of views in datasets without having to discard any available measurement.
Moreover, the common latent representation inferred with our model can be used to define robust classification tasks gathering the combined information across different datasets.
